Convective penetration probably parameterizes convective overshoot

### 2021-07-03 manuscript/ms.tex
adds some changes to theory section of ms.

Eqn.~\ref{eqn:integral_constraint} is the simple expression of \citet{roxburgh1989}'s integral constraint (is it?), and is the straightforward statement that work by buoyancy in the convection zone on large scales must be balanced by viscous dissipation on small scales.

@@

We see that the size of the PZ is set by some combination of buoyancy breaking and viscous dissipation, whose total effect cancels out any remaining work that dissipation in the convection zone did not itself cancel.
we now find it instructive to define the fraction of the buoyant engine consumed by dissipation in the convection zone

@@

f \equiv \frac{\int_{\rm{CZ}} \bar{\Phi}\,dz}{\int_{\rm{CZ}}\bar{\mathcal{B}}\,dz},

@@

and note two interesting limits,

@@

It is probably also reasonable to assume that $\int_{\rm{PZ}}\bar{\Phi}\,dz \rightarrow 0$, and so Eqn.~\ref{eqn:constraint_cz_pz_split} states that the PZ must be large enough to cancel out all of the buoyant work done by convection in the CZ by buoyancy breaking in the PZ.

@@

\item In the limit that $f \rightarrow 1$, viscous dissipation is extremely efficient in the convection zone and cancels out all of the work done by buoyancy.

@@

Per Eqn.~\ref{eqn:constraint_cz_pz_split}, the positive definite PZ terms must approach zero and no PZ develops in this limit.

@@

Indeed, we find that $f \gg 0$ but $f < 1$ in our simulations (see Fig.~\ref{fig:theory_profiles}, right panel).

In order to say more about the size of the penetration layer, we model $\bar{\Phi}$ as a linearly decreasing function of height in the PZ.
We sets its value at the top of the PZ equal to the average of $\bar{\Phi}$ in the CZ of depth $L_{\rm{CZ}}$ so that $\int_{\rm{PZ}} \bar{\Phi}\,dz = (\delp/[2 L_{\rm{CZ}}])\int_{\rm{CZ}}\bar{\Phi}\,dz$ where $\delp$ is the penetration zone depth.
Rearranging Eqn.~\ref{eqn:constraint_cz_pz_split} and dividing by $\int_{\rm{CZ}} \bar{\mathcal{B}}\,dz$, and plugging in $\bar{\mathcal{B}} = |\alpha|g\bar{\Fconv}$, we can write
\begin{equation}
-\frac{\int_{\rm{PZ}}\bar{\Fconv}\,dz}{\int_{\rm{CZ}}\bar{\Fconv}\,dz} + f\frac{\delp}{2 L_{\rm{CZ}}}
\approx (1 - f).
\label{eqn:theory_fraction}
\end{equation}
In order to derive a specific prediction for the PZ depth, it is necessary to specify the vertical shape of $\overline{\Fconv}$.
We will study two cases in this work, laid out below.
In both of these cases, we will define a nondimensional ``Penetration Parameter'' whose magnitude is set by the ratio of the convective flux some equidistant length $\epsilon$ both above and below the Schwarzschild convective boundary $L_s$ (assuming the region above the boundary is an adiabatic penetrative zone),
\begin{equation}
\mP \equiv -\frac{\overline{\Fconv}(z = L_s - \epsilon)}{\overline{\Fconv}(z = L_s + \epsilon)}.
\label{eqn:theory_P_defn}
\end{equation}
Since $\Fconv < 0$ in the PZ, the sign of $\mP$ is positive.

@@

\subsubsection{Case I: Discontinuous flux}
\label{sec:discontinuous_theory}
We first consider a model which satisfies
\begin{equation}
\overline{\Fconv}(z) = \Fcz \begin{cases}
1			&	z \leq \Lcz,\\
-\mP_D^{-1}  & 	z > \Lcz 
\end{cases}.
\end{equation}
Here, $\Fcz$ is a constant value of flux carried in the convection zone and $\mP_D$ is the penetration parameter (subscript D for discontinuous case).
Plugging this functional form of the flux into Eqn.~\ref{eqn:theory_fraction}, and integrating the CZ over a depth $L_{\rm{CZ}}$ below $L_s$ and the PZ over a depth $\delp$ above $L_s$, we predict
\begin{equation}
\frac{\delp}{L_{\rm{CZ}}} \approx \mP_D \frac{1 - f}{1 + \mP_D f / 2}.
\label{eqn:discontinuous_prediction}
\end{equation}
Assuming that $f$ is a weak function of $\mP_D$, we see that the size of the penetration region is roughly linearly proportional to $\mP_D$ at small $\mP_D$, but saturates due to dissipation at large $\mP_D$.
Intuitively, this result makes sense: as $\mP_D$ grows, the magnitude of $\overline{\Fconv}$ and the breaking force of buoyancy in the PZ shrink, resulting in larger penetrative regions (but this growth cannot extend indefinitely).

\subsubsection{Case II: Piecewise linear flux}
\label{sec:linear_theory}
We next assume that $\overline{\Fconv}(z)$ is not discontinuous at the CZ-PZ boundary, but that its derivative may be,
\begin{equation}
\overline{\Fconv}(z) = 
\frac{\partial \Frad}{\partial z}\bigg|_{\rm{CZ}}
\begin{cases}
(L_p - z) & z \leq L_p \\
-\mP_L^{-1} (z - L_p) & z > L_p
\end{cases},
\end{equation}
where $(\partial \Frad / \partial z)|_{\rm{CZ}}$ is a constant and $\mP_L$ is the penetration parameter (subscript L for linear case).
Again, solving Eqn.~\ref{eqn:theory_fraction} with this functional form of the flux and integrating over $L_{\rm{CZ}}$ in the CZ and $\delp$ in the PZ, we retrieve a quadratic equation.
This equation has two solution branches, only one of which corresponds to a positive value of $\delp$.
On that branch, we find
\begin{equation}
\frac{\delp}{L_p} = \sqrt{\mP_L (1 - f)} \,\,(\sqrt{\xi^2 + 1} - \xi),
\label{eqn:linear_prediction}
\end{equation}
where $\xi \equiv (f/4)\sqrt{\mP_L/(1-f)}$.
So we expect the penetration depth to be roughly proportional to $\sqrt{\mP_L}$ for small values of $\mP_L$, and to again saturate at large values of $\mP_L$.

@@

In our equations, we expect the volume-average convective velocities to depend on the magnitude of the heating, $\angles{\vec{u}} \approx Q_{\rm{mag}}^{1/2} \approx 1$, so the characteristic convective frequency is $f_{\rm{conv}} \approx \angles{\vec{u}} L_s \approx 1$.
The stiffness is defined,
\begin{equation}
\mS \equiv \frac{N^2}{f_{\rm{conv}}} \approx N^2,
\end{equation}
where $N^2$ is the \brunt$\,$frequency in the radiative zone.
In our nondimensionalization, $N^2 = \gradad - \gradrad$ where $\gradrad = \Ftot/k$.
We treat $\mS$ as a control parameter; by choosing a value of the stiffness we set the magnitude of the background temperature gradient, $\partial_z T_0$ which in turn sets the value of $k$ in the radiative zone.

>>>>

Integral constraints of this form are the basis for a broad range of analyses in Boussinesq convection \citep[see e.g.,][]{ahlers_etal_2009, goluskin2016} and were considered in the context of penetrative convection by \citet{roxburgh1989}.
Eqn.~\ref{eqn:integral_constraint} is the straightforward statement that work by buoyancy on large scales must be balanced by viscous dissipation on small scales.

@@

We see that the size of the PZ is set by some combination of viscous dissipation and buoyancy breaking, whose total effect cancels out any remaining work that dissipation in the convection zone did not itself cancel.
we now find it instructive to parameterize the fraction of the buoyant engine consumed by CZ dissipation  

@@

f \equiv \frac{\int_{\rm{CZ}} \bar{\Phi}\,dz}{\int_{\rm{CZ}}\bar{\mathcal{B}}\,dz}.

@@

Under this parameterization, Eqn.~\ref{eqn:constraint_cz_pz_split} can be written
\frac{\int_{\rm{PZ}} \bar{\Phi}\,dz + \int_{\rm{PZ}}(-\bar{\mathcal{B}})\,dz}{\int_{\rm{CZ}} \bar{\mathcal{B}}\,dz}
= 1 - f,
\label{eqn:first_pz_parameterization}
and we can immediately read off limits on a hypothetical PZ according to $f$,

@@

It is probably also reasonable to assume that $\int_{\rm{PZ}}\bar{\Phi}\,dz \rightarrow 0$, and so Eqn.~\ref{eqn:first_pz_parameterization} states that the PZ must be large enough for its negative buoyant work to be equal in magnitude to the positive buoyant work of the CZ.

@@

\item In the limit that $f \rightarrow 1$, viscous dissipation is efficient in the CZ and cancels out the work done by buoyancy.

@@

Per Eqn.~\ref{eqn:first_pz_parameterization}, the positive-definite PZ terms must approach zero and no PZ develops in this limit.

@@

Indeed, we find that $f \gg 0$ but $f < 1$ in our simulations (see e.g., Fig.~\ref{fig:theory_profiles}, right panel).

Assuming that a PZ of depth $\delp$ develops above a CZ of depth $L_{\rm{CZ}}$, we model the PZ dissipation as
\begin{equation}
\int_{\rm{PZ}} \bar{\Phi}\,dz = \xi\frac{\delp}{L_{\rm{CZ}}}\int_{\rm{CZ}}\bar{\Phi} = \xi \delp \Phi_{\rm{CZ}}.
\end{equation}
Here $\Phi_{\rm{CZ}}$ is the average dissipation value in the CZ and $\xi$ is a parameter that describes the shape of the dissipation profile as a function of height in the PZ.
In other words, we assume that $\bar{\Phi} \approx \Phi_{\rm{CZ}}$ at the CZ-PZ boundary and then falls off with height.
The shape of $\bar{\Phi}$ determines the value of $\xi$, with a linear falloff producing $\xi = 1/2$, a quadratic falloff producing $\xi = 2/3$, and the in which $\bar{\Phi}$ does not decline with height produces $\xi = 1$.
Regardless, $\xi$ is another measurable parameter bounded by $[0, 1]$. 
With this parameterization, we rewrite Eqn.~\ref{eqn:first_pz_parameterization}
\begin{equation}
-\frac{\int_{\rm{PZ}}\bar{\Fconv}\,dz}{\int_{\rm{CZ}}\bar{\Fconv}\,dz} + f\xi\frac{\delp}{L_{\rm{CZ}}}
= (1 - f).
\label{eqn:theory_fraction}
\end{equation}
In order to derive a specific prediction for the PZ depth, it is necessary to specify the vertical shape of $\overline{\Fconv}$.
We will study two cases in this work, laid out below.
In both of these cases, we will define a nondimensional ``Penetration Parameter'' whose magnitude is set by the ratio of the convective flux some equidistant length $\epsilon$ both above and below the Schwarzschild convective boundary $L_s$ (assuming the region above the boundary is an adiabatic penetrative zone),
\begin{equation}
\mP \equiv -\frac{\overline{\Fconv}(z = L_s - \epsilon)}{\overline{\Fconv}(z = L_s + \epsilon)}.
\label{eqn:theory_P_defn}
\end{equation}
Since $\Fconv < 0$ in the PZ, the sign of $\mP$ is positive.

@@

\subsubsection{Case I: Discontinuous flux}
\label{sec:discontinuous_theory}
We first consider a model which satisfies
\begin{equation}
\overline{\Fconv}(z) = \Fcz \begin{cases}
1			&	z \leq \Lcz,\\
-\mP_D^{-1}  & 	z > \Lcz 
\end{cases}.
\end{equation}
Here, $\Fcz$ is a constant value of flux carried in the convection zone and $\mP_D$ is the penetration parameter (subscript D for discontinuous case).
Plugging this functional form of the flux into Eqn.~\ref{eqn:theory_fraction}, and integrating the CZ over a depth $L_{\rm{CZ}}$ below $L_s$ and the PZ over a depth $\delp$ above $L_s$, we predict
\begin{equation}
\frac{\delp}{L_{\rm{CZ}}} = \mP_D \frac{1 - f}{1 + \xi f \mP_D}.
\label{eqn:discontinuous_prediction}
\end{equation}
Assuming that $f$ is a weak function of $\mP_D$, we see that the size of the penetration region is roughly linearly proportional to $\mP_D$ at small $\mP_D$, but saturates due to dissipation at large $\mP_D$.
Intuitively, this result makes sense: as $\mP_D$ grows, the magnitude of $\overline{\Fconv}$ and the breaking force of buoyancy in the PZ shrink, resulting in larger penetrative regions (but this growth cannot extend indefinitely).

\subsubsection{Case II: Piecewise linear flux}
\label{sec:linear_theory}
We next assume that $\overline{\Fconv}(z)$ is not discontinuous at the CZ-PZ boundary, but that its derivative may be,
\begin{equation}
\overline{\Fconv}(z) = 
\frac{\partial \Frad}{\partial z}\bigg|_{\rm{CZ}}
\begin{cases}
(L_p - z) & z \leq L_p \\
-\mP_L^{-1} (z - L_p) & z > L_p
\end{cases},
\end{equation}
where $(\partial \Frad / \partial z)|_{\rm{CZ}}$ is a constant and $\mP_L$ is the penetration parameter (subscript L for linear case).
Again, solving Eqn.~\ref{eqn:theory_fraction} with this functional form of the flux and integrating over $L_{\rm{CZ}}$ in the CZ and $\delp$ in the PZ, we retrieve a quadratic equation.
This equation has two solution branches, only one of which corresponds to a positive value of $\delp$.
On that branch, we find
\begin{equation}
\frac{\delp}{L_p} = \sqrt{\mP_L (1 - f)} \,\,(\sqrt{\zeta^2 + 1} - \zeta),
\label{eqn:linear_prediction}
\end{equation}
where $\zeta \equiv (\xi f/2)\sqrt{\mP_L/(1-f)}$.
So we expect the penetration depth to be roughly proportional to $\sqrt{\mP_L}$ for small values of $\mP_L$, and to again saturate at large values of $\mP_L$.

@@

In our equations, we expect the volume-average convective velocities to depend on the magnitude of the heating, $\angles{\vec{u}} \approx Q_{\rm{mag}}^{1/2} \approx 1$, so the characteristic convective frequency is $f_{\rm{conv}} \approx \angles{\vec{u}} L_s \approx 1$.
The stiffness is defined,
\begin{equation}
\mS \equiv \frac{N^2}{f_{\rm{conv}}^2} \approx N^2,
\end{equation}
where $N^2$ is the \brunt$\,$frequency in the radiative zone.
In our nondimensionalization, $N^2 = \gradad - \gradrad$ where $\gradrad = \Ftot/k$.
We treat $\mS$ as a control parameter; by choosing a value of the stiffness we set the magnitude of the background temperature gradient, $\partial_z T_0$ which in turn sets the value of $k$ in the radiative zone.
